% HW 4: Fitting Accident Data with regression line 
% Robert Smith
% \today

```{r setup, include=FALSE}
library(forecast)
library(tseries)
library(gridExtra)
library(ggplot2)
library(scales)

opts_knit$set(fig.path="figure/hw4-",fig.height=8,fig.width=7.5)

#--Produces a data.frame with the Source Data+Training Data, Fitted Values+Forecast Values, forecast data Confidence Intervals
funggcast <- function(dn, fcst){ 
  require(zoo) #needed for the 'as.yearmon()' function
  
  en <- max(time(fcst$mean)) #extract the max date used in the forecast
  
  #Extract Source and Training Data
  ds <- as.data.frame(window(dn, end = en))
  names(ds) <- 'observed'
  ds$date <- as.Date(time(window(dn, end = en)))
  
  #Extract the Fitted Values (need to figure out how to grab confidence intervals)
  dfit <- as.data.frame(fcst$fitted)
  dfit$date <- as.Date(time(fcst$fitted))
  names(dfit)[1] <- 'fitted'
  
  ds <- merge(ds, dfit, all.x = T) #Merge fitted values with source and training data
  
  #Exract the Forecast values and confidence intervals
  dfcstn <- as.data.frame(fcst)
  dfcstn$date <- as.Date(as.yearmon(row.names(dfcstn)))
  names(dfcstn) <- c('forecast', 'lo80', 'hi80', 'lo95', 'hi95', 'date')
  
  pd <- merge(ds, dfcstn, all.x = T) #final data.frame for use in ggplot
  return(pd)
  
}

plot_fcst <- function(dn, fcst) {
  pd <- funggcast(dn, fcst)
  
  return( 
    ggplot(data = pd, aes(x = date, y = observed)) + 
      geom_ribbon(aes(ymin = lo95, ymax = hi95), 
                  fill = 'orange', alpha =  0.25) + 
      geom_ribbon(aes(ymin = lo80, ymax = hi80), 
                  fill = 'orange', alpha = 0.5) + 
      geom_line(col = 'red') + 
      geom_line(aes(y = fitted), col = 'blue') + 
      geom_line(aes(y = forecast)) 
    )
  
}

```

#### 1. Read in acci.txt file from the course web site. Plot the time series. Take difference with lag 12. Plot the differenced data. Call it D2.

```{r q1,echo=c(1:3,10:13),warning=FALSE,error=FALSE}
acci <- read.table("acci.txt", header=TRUE)
D1 <- ts(acci, frequency=12)
D2 <- diff(D1, d = 12)

# layout(matrix(c(1,1,2,3), 2, 2, byrow = TRUE))
# plot(D1, main = "ACCI Time Series", xlab = "time", ylab = "value")
# acf(D1, main = "ACCI Time Series ACF", xlab = "lag")
# pacf(D1, main = "ACCI Time Series PACF", xlab = "lag")

adf.test(D2)
pp.test(D2)
kpss.test(D2)

layout(matrix(c(1,1,2,3), 2, 2, byrow = TRUE))
plot(D2, main = "ACCI Time Series with Year-over-Year (YOY) differencing", xlab = "time", ylab = "value")
acf(D2, main = "ACCI YOY ACF", xlab = "lag")
pacf(D2, main = "ACCI YOY PACF", xlab = "lag")

# layout(matrix(c(1,2), 2, 1, byrow = TRUE))
# plot(D1)
# 
# acci_df <- data.frame(
#   date = 1:nrow(acci), 
#   values = acci$d,
#   D12 = c(rep(NA,12), as.numeric(D2))
#   )
# 
# pD1 <- ggplot(acci_df, aes(date, values)) + geom_line(stat="identity") + scale_y_continuous(labels = comma) + scale_x_continuous(breaks = c(12,24,36,48,60,72)) + ggtitle("ACCI Values")
# 
# pD2 <- ggplot(acci_df, aes(date, D12)) + geom_line(stat="identity") + scale_y_continuous(labels = comma) + scale_x_continuous(breaks = c(12,24,36,48,60,72)) + ggtitle("ACCI Values - Year over Year Differencing")
# 
# grid.arrange( pD1, pD2, ncol = 1 )
```

#### 2. Fit D2 with linear trend using OLS. Is the slope significant? How did you determine? Can the output from the screen be trusted?    



Based on the RSE, adjusted $\rho^2$ and p-value we fail to reject $H_0$ and cannot regard the slope of the OLS fit of the year-over-year differences as signficant. This indicates that the values obtained in D2 are dependent upon time and should probably not be trusted. When we plot the forecast of the residuals we see a repeating pattern in the data and the forecast as well.

```{r q2,echo=1:17,warning=FALSE,error=FALSE}
time <- 1:length(D2)
(err_mod <- summary(lm_D2 <- lm(D2~time)))
resid <- lm_D2$residuals

(arima_mdl <- auto.arima(ts(resid, start = c(2,1), freq = 12)))

adf.test(resid)
pp.test(resid)

layout(matrix(c(1, 2, 3), 3, 1, byrow = TRUE))
plot(D2); abline(lm_D2)
plot(x = 1:length(resid), y = resid, type = "l", main = "Fits vs. Residuals", ylab = "residuals")
plot(forecast(arima_mdl))

layout(matrix(c(1,2,3,4), 2, 2, byrow = TRUE))
plot(lm_D2)

# D12_res <- data.frame(
#   date = 1:nrow(acci),
#   residuals = c(rep(NA, 12), lm_D2$residuals)
#   )
# 
# ggplot(D12_res, aes(x = date, y = residuals)) + 
#   geom_line(stat="identity") + 
#   scale_y_continuous(labels = comma) + 
#   scale_x_continuous(breaks = c(12,24,36,48,60,72)) + 
#   ggtitle("ACCI D12 OLS Residuals")

```

#### 3. Fit residuals from (2) with seasonal ARIMA with d=0, D=0. (i.e. ARIMA(p,0,q)x(P,0,Q)12). Is the seasonal part necessary?

Based on the results above, the standard error for the seasonal component is significant and therefore should be included in the forecast.


```{r q3,warning=FALSE,error=FALSE}

(arima00 <- auto.arima(
  ts(lm_D2$residuals, 
     start = c(2,1), 
     freq = 12), 
  d=0, 
  D=0))

```

#### 4. Using the best model from #3, predict twelve months ahead in D2.

```{r q4,warning=FALSE,error=FALSE}

plot(forecast(arima00, h = 12))

```

#### 5. Using the prediction from #4, predict twelve months ahead in D1 (original TS).

```{r q5,warning=FALSE,error=FALSE}

(q5D1 <- Arima(
  D1, 
  order = c(3,0,0), 
  seasonal = list(
    order = c(1,0,0), 
    period = 12)
  ))

q5 <- forecast(q5D1)
plot(q5)

```

#### 6. Using your model from In-class Ex2-#2, (ARIMA(p,1,q)x(P,1,Q)12 model), predict twelve months ahead in D1.

```{r q6,warning=FALSE,error=FALSE}

(q6D1 <- auto.arima(D1, d = 1, D = 1))
q6 <- forecast(q6D1)
plot(q6)

```

#### 7. Using your model from In-class Ex2-#3, (ARIMA(p,0,q)x(P,1,Q)12 model), predict twelve months ahead in D1.

```{r q7,warning=FALSE,error=FALSE}

(q7D1 <- auto.arima(D1, d = 0, D = 1))
q7 <- forecast(q7D1)
plot(q7)

```

#### 8. Compare your prediction from #5, #6, and #7. Plot the one-month predictions on the sampe plot. Which one do you like the best?

Based on the results from questions 5, 6 & 7 I believe the model fit in question #7 to be the best based on its AIC, AICc & BIC statistics being the smallest for the three questions and for the standard errors it produces being among the smallest of the group.

```{r q8,warning=FALSE,error=FALSE}

plot(D1, xlim = c(1,8))
points(x = seq(7,8, length.out=12), y = q7$mean[1:12])
points(x = seq(7,8, length.out=12), y = q6$mean[1:12], col = "red")
points(x =seq(7,8, length.out=12), y = q5$mean[1:12], col = "blue")

```


